Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-19 cordillera-cycle.tex
Small language improvements.

  Despite more than a century of geological observations, the Cordilleran ice
  sheet of North America remains poorly understood in terms of its former
  extent, volume and dynamics. Although geomorphological evidence is abundant,
  its complexity is such that whole ice-sheet reconstructions of advance and
  retreat patterns are lacking. Here we use a numerical ice sheet model
  calibrated against field-based evidence to attempt a quantitative
  reconstruction of the Cordilleran ice sheet history through the last glacial
  cycle. A series of simulations is driven by time-dependent temperature
  offsets from six proxy records located around the globe. Although this
  approach reveals large variations in model response to evolving atmospheric
  forcing, all simulations produce two major glaciation events during
  marine isotope stages~4 (61.9--55.4\,kyr) and~2
  (29.5--16.9\,kyr). The timing of glaciation is
  better reproduced using temperature reconstructions from Greenland and
  Antarctic ice cores than from regional oceanic sediment cores. During most of
  the last glacial cycle, the modelled ice cover is discontinuous and
  restricted to high mountain areas. However, widespread precipitation over the
  Skeena mountains favours the persistence of a central ice dome throughout the
  glacial cycle. It acts as a nucleation centre before the last glacial maximum
  and hosts the last remains of Cordilleran ice during the
  early Holocene (10.9--9.5\,kyr).

@@

The present study uses a numerical ice sheet model \citep{PISM-authors.2014},
calibrated against field-based evidence to perform a quantitative
reconstruction of the Cordilleran ice sheet history through the last glacial
cycle. Although numerical modelling has been established as a useful tool to
improve our understanding of the Cordilleran ice sheet more than twenty years
ago (\citealp[p.~227]{Jackson.Clague.1991}; \citealp{Robert.1991}), the
ubiquitously mountainous
topography of the region has presented a major challenge to its application.
In fact, only recent development of numerical ice sheet models and underlying
scientific computing tools \citep{Bueler.Brown.2009, Balay.etal.2014} allows
high-resolution numerical modelling of palaeo-glaciers on mountainous terrain
over millenial time scales \citep{Golledge.etal.2012}. In the North
American Cordillera, this complex topography induces strong gradients in
seasonality and distinct patterns of precipition, thus requiring the use of
high-resolution atmospheric forcing fields as input to the ice sheet model
\citep{Seguinot.etal.2013}.

>>>>

  Despite more than a century of geological observations, the Cordilleran ice
  sheet of North America remains poorly understood in terms of its former
  extent, volume and dynamics. Although geomorphological evidence is abundant,
  its complexity is such that whole ice-sheet reconstructions of advance and
  retreat patterns are lacking. Here we use a numerical ice sheet model
  calibrated against field-based evidence to attempt a quantitative
  reconstruction of the Cordilleran ice sheet history through the last glacial
  cycle. A series of simulations is driven by time-dependent temperature
  offsets from six proxy records located around the globe. Although this
  approach reveals large variations in model response to evolving atmospheric
  forcing, all simulations produce two major glaciation events during
  marine isotope stages~4 (61.9--55.4\,kyr) and~2
  (29.5--16.9\,kyr). The timing of glaciation is
  better reproduced using temperature reconstructions from Greenland and
  Antarctic ice cores than from regional oceanic sediment cores. During most of
  the last glacial cycle, the modelled ice cover is discontinuous and
  restricted to high mountain areas. However, widespread precipitation over the
  Skeena Mountains favours the persistence of a central ice dome throughout the
  glacial cycle. It acts as a nucleation centre before the last glacial maximum
  and hosts the last remains of Cordilleran ice during the
  early Holocene (10.9--9.5\,kyr).

@@

The present study uses a numerical ice sheet model \citep{PISM-authors.2014},
calibrated against field-based evidence to perform a quantitative
reconstruction of the Cordilleran ice sheet history through the last glacial
cycle. Although numerical modelling has been established as a useful tool to
improve our understanding of the Cordilleran ice sheet more than twenty years
ago (\citealp[p.~227]{Jackson.Clague.1991}; \citealp{Robert.1991}), the
ubiquitously mountainous
topography of the region has presented a major challenge to its application.
In fact, only recent development of numerical ice sheet models and underlying
scientific computing tools \citep{Bueler.Brown.2009, Balay.etal.2014} allows
high-resolution modelling of palaeo-glaciers on mountainous terrain
over millenial time scales \citep{Golledge.etal.2012}. In addition, complex
topography of the North American Cordillera induces strong gradients in
seasonality and distinct patterns of precipition, thus requiring the use of
high-resolution atmospheric forcing fields as input to the ice sheet model
\citep{Seguinot.etal.2013}.
